`{peacesciencer}`: An R Package for Quantitative Peace Science Research

### 2022-03-22 manuscript/ms.Rmd
add, with professional easter egg

---
title: "`{peacesciencer}`: An R Package for Quantitative Peace Science Research"
thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
author:
- name: Steven V. Miller
  affiliation: Clemson University
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
# bibliography: inst/refs.bib
abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate common types of data from scratch and to merge in ubiquitous indicators included in almost every analysis (e.g. democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at `r params$ps_website`."
keywords: "software, statistical analysis, peace science, reproducibility"
geometry: margin=1in
csl: inst/jpr.csl
mainfont: cochineal
monofont: Fira Code
#sansfont: Linux Biolinum O 
fontsize: 11pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
  ps_website: ""
  ps_manuals: ""
  ps_github: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
nocite: |
  @cartersmith2020fml, @marshalletal2017p, @coppedgeetal2020vdem, @gleditschetal2002ac, @thompsondreyer2012hir, @cowstates2016, @stinnettetal2002cow, @pevehouseetal2020tow, @schvitz2021mis, @palmeretal2021mid5, @singer1987rcwd, @nardulli2012creg, @gledtischward1999rlis, @nunnpuga2012r, @haege2011cc, @pettersson2019ov, @leedsetal2002atop, @barbierietal2009td
#always_allow_html: TRUE
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \usepackage{csquotes}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
#  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
#  - \usepackage{colortbl}
#  - \arrayrulecolor{black}
#  - \usepackage{threeparttable}
#  - \usepackage{threeparttablex}
#  - \midrule\relax
---

@@

`{peacesciencer}` is designed to address these problems. Built around the free and open source R programming language, `{peacesciencer}` contains a suite of data and functions for creating data of interest to researchers. Researchers can use `{peacesciencer}` to create dyad-year, leader-year, leader-dyad-year, and state-year data (among some others) from scratch. Afterward, they can add a variety of standard information (e.g. contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. `{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing conflicts. All are done with the maximum possible transparency. The project is available for public view on Github (`r params$ps_github`). The `data-raw` directory on the project's Github contains information and comments about how every data set was created. The function manuals (`r params$ps_manual`) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, `{peacesciencer}` not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

@@

`{peacesciencer}` is motivated by the following observations and ideals that led to its creation. For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g. dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. `{peacesciencer}` is primarily born from this question about how to create these data from scratch. The underlying code that produces these data types is available online and `{peacesciencer}` converts these lines of code into simple functions for the ease of the researcher.

@@

tribble(~func, ~type, ~description,
        "add_archigos()", "D,S", "Add Archigos political leader information to a data frame",
        "add_atop_alliance()", "D, LD", "Add Alliance Treaty Obligations and Provisions (ATOP) alliance data to a data frame",
        "add_capital_distance()", "D, L, LD, S", "Add capital-to-capital distance to a data frame",
        "add_ccode_to_gw()", "D, L, LD, S", "Match CoW state codes to G-W state codes",
        "add_contiguity()", "D, L, LD, S", "Add CoW direct contiguity information to a data frame",
        "add_cow_alliance()", "D, LD", "Add CoW alliance data to a data frame",
        "add_cow_majors()", "D, L, LD, S", "Add CoW major power information to a data frame",
        "add_cow_mids()", "D", "Add CoW-MID data to a dyad-year data frame",
        "add_cow_trade()", "D, L, LD, S", "Add CoW trade data to a data frame",
        "add_cow_wars()", "D, S", "Add CoW war data to a dyad-year or state-year data frame",
        "add_creg_fractionalization()", "D, L, LD, S", "Add fractionalization/polarization estimates from CREG to a data frame",
        "add_democracy()", "D, L, LD, S", "Add democracy information to a data frame",
        "add_fpsim()", "D, LD", "Add dyadic foreign policy similarity measures to a data frame",
        "add_gml_mids()", "D, L, LD, S", "Add GML MID data to a data frame",
        "add_gwcode_to_cow()", "D, L, LD, S", "Match G-W state codes to CoW state codes",
        "add_igos()", "D, S", "Add CoW IGO data to a data frame",
        "add_lead()", "L, LD", "Add leader experience/attributes (LEAD) to data",
        "add_lwuf()", "L, LD", "Add leader willingness to use force estimates to data",
        "add_minimum_distance()", "D, L, LD, S", "Add minimum distance data to data frame",
        "add_nmc()", "D, L, LD, S", "Add Correlates of War National Military Capabilities data",
        "add_peace_years()", "D, S", "Add 'peace years' to dyad-year/state-year conflict data",
        "add_rugged_terrain()", "D, L, LD, S", "Add rugged terrain information to a data frame",
        "add_sdp_gdp()", "D, L, LD, S", "Add (surplus, gross) domestic product data to a data frame",
        "add_spells()", "D, L, LD, S", "Add duration 'spells' to a data frame",
        "add_strategic_rivalries()", "D, S", "Add strategic rivalry information to a data frame",
        "add_ucdp_acd()", "S", "Add UCDP Armed Conflict Data to state-year data frame",
        "add_ucdp_onsets", "S", "Add UCDP onsets to state-year data",
        "filter_prd()", "D, LD", "Filter dyad-year data to just politically relevant dyads",
        "wc_duration()", "C", "Whittle Duplicate Conflict-Years by (Minimum or Maximum) Conflict Duration",
        "wc_fatality()", "C", "Whittle Duplicate Conflict-Years by Highest Fatality",
        "wc_hostility()", "C", "Whittle Duplicate Conflict-Years by Highest Hostility",
        "wc_jds()", "C", "Whittle Duplicate Conflict-Years by Just Dropping Something",
        "wc_onsets()", "C", "Whittle Uniqute Conflict Onset-Years from Conflict-Year Data",
        "wc_recip()", "C", "Whittle Duplicate Conflict-Years by Conflict Reciprocation",
        "wc_stmon()", "C", "Whittle Duplicate Conflict-Years by Lowest Start Month"
) -> add_functions

@@

`{peacesciencer}` is a package for the R programming language. This assumes at least some familiarity with the R programming language. Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.2. `{peacesciencer}` is designed to be as user-friendly as possible. Those proficient in R, those just learning R, and those with no experience in R should be able to pick up its use fairly quickly.

@@

[^poundsign]: R users will recognize that pound signs (`#`) allow for commenting code. Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing. In subsequent code snippets, `#>` will indicate console output the user can expect from these functions.

@@

I encourage users who are using `{peacesciencer}` for the first time, especially those who are learning R for the first time because of their interest in this package, to approach `{peacesciencer}` with an idea of the kind of data they want to create for the sake of a project. The core functionality of `{peacesciencer}` begins with the creation of a data frame, which can then be populated with various indicators of interest. No matter, the suggested use of `{peacesciencer}` begins with the creation of a data frame. To start, assume a new user without much familiarity with the R programming language installed RStudio, `{tidyverse}`, and `{peacesciencer}` with the idea of using `{peacesciencer}` to help them start a new research project that seeks to explain civil conflict onset across state-years. Toward that end, they have identified their unit of analysis is state-year and can be created with the `create_stateyears()`. To get started, I encourage entering the following command in the console in RStudio.

@@

A user who approaches `{peacesciencer}` with a project in mind will see they can better tailor this function to what they want. Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the UCDP armed conflict data. They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward. Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).

@@

This function has three arguments denoting the leader system to inform the creation of the leader-year data (i.e. Archigos) and two other arguments: `standardize` and `subset_years`. The `standardize` argument, which when it defaults to "none", returns all leader-years as presented in the raw Archigos data. Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g. Archigos often has leader entries prior to state system entry in a few cases). Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates of CoW state system dates. Finally, the user can subset the leader-year data to a more narrow temporal domain that may interest them with the `subset_years` function. If a user is interested in creating a data set of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function. Notice how `create_leaderyears()` also returns information about the leader too, like the leader's approximate age that year, their gender, and information about their tenure.

@@

Researchers using `{peacesciencer}` to create data for their research project should start with one of these "create" functions. Whether state-year, dyad-year, or a leader-level analysis, these functions will create the full universe of cases of interest to a researcher for the unit of analysis that serves as the core of their project.

@@

Researchers should see that `{peacesciencer}`'s functionality can scale up nicely from there. For example, the following would round out the kind of information necessary to replicate Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis by adding information about national material capabilities for both states in the dyad (`add_nmc()`), estimates of democracy for both states in the dyad (`add_democracy()`), information about alliance commitments in the dyad-year (`add_cow_alliance()`, by way of @gibler2009ima), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by @andersetal2020bbgb. Whereas `add_sdp_gdp()` is the last command in the pipe-based workflow, the `{peacesciencer}` call ends by assigning to an object called `Data`. This type of assignment is done with the "right hand" assignment operator (i.e. `->`).

@@

If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like `haven::write_dta(Data, "my-data.dta")` and import it into Stata when they are done. No matter, `{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that has greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the capabilities data. In this particular application, it has already created the main data required for a replication of @bremer1992ddc. There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land-contiguity, the presence of a major power in the dyad, and some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme method for extending the Unified Democracy Scores data), and the GDP per capita in the dyad. Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators. Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but `{peacesciencer}` greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.

@@

`{peacesciencer}` is capable of generating data for replications of analyses at multiple levels beyond dyad-year. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with `create_stateyears(system = 'gw', subset_years = c(1946:2019))`, creating the full universe of Gleditsch-Ward state years and subsetting them to just 1946-2019 (because the UCDP data included in `{peacesciencer}` include just the observations in that time frame). Next, we can use the `add_ucdp_acd()` function to return information about ongoing UCDP conflicts and onsets for these states. `add_ucdp_acd()` takes three arguments: `type`, `issue`, and `only_wars`. `type` is an optional argument for the type of armed conflicts for which the researcher wants information. Options include "extrasystemic", "interstate", "intrastate", and "II" (short for "internationalized intrastate"). If no `type` is specified, the function returns information about ongoing disputes and onsets for all states for all types of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g. `type = c("intrastate", "II")`). `issue` is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include "territory", "government", and "both". If no `issue` is specified, the function returns information for all conflicts regardless of the particular issue. `only_wars` is an argument that subsets the data to just those with the intensity levels of "war" when `only_wars = TRUE`. The argument defaults to `FALSE`, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, `add_ucdp_acd(type = "intrastate", only_wars = FALSE))` returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.[^adducdpacdalso]

@@

Finally, we can add some covariates of interest to these data. `add_spells()` calculates peace spells between ongoing conflicts in the data generated by `add_ucdp_acd()`. `add_democracy()` adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' [-@marquez2016qme] extension of Pemstein et al's [-@pemsteinetal2010dc] Unified Democracy Scores). `add_creg_fractionalization()` adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. `add_sdp_gdp()` will add information about a state's estimated GDP, population, and GDP per capita from the @andersetal2020bbgb simulations. Finally, `add_rugged_terrain()` provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by @nunnpuga2012r and the second is the @giblermiller2014etts extension of the natural logged percentage of the state that is mountainous (originally calculated by @fearonlaitin2003eicw). At the end of the pipe, the data returned by `{peacesciencer}` is assigned to an object minimally called `Data`.

`{peacesciencer}`'s tight integration with the `{tidyverse}` permits wide flexibility for the researcher. For example, assume the researcher wants to discern the estimated effect of the same set of covariates on intra-state conflicts at the threshold of war and those intra-state conflicts at or below the threshold of war. The first call included all conflicts with at least 25 deaths, per the UCDP's inclusion rules, and the peace years were calculated for those as well. If the researcher wants a new set of conflicts with a new set of peace years, it would be a simple matter of repeating the pipe-based workflow, but altering the argument in `add_ucdp_acd()` to be `only_wars = TRUE`. `{peacesciencer}` would then calculate the peace years for those (`add_peace_years()`). To avoid confusion with the overlapping column names, the researcher can use some `{tidyverse}` verbs to rename all those conflict variables to have a distinct prefix of `war_` (i.e. `rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .))`) before finally joining these data into the master data frame (i.e. `left_join(Data, .) -> Data`).  Table \@ref(tab:tab-cw) shows the fruits of the data `{peacesciencer}` generated after some post-processing and lagging important variables.

@@

`{peacesciencer}` also has support for newer levels of analysis in the peace science community, prominently leader-levels of analysis. There has been considerable emphasis in peace science research to emphasize that state leaders, not "states", make foreign policy decisions that may lead to war. Understanding the attributes of leaders themselves are critical to the core research questions of the community [@ellisetal2015lead; @goemansetal2009ia; @horowitzstam2014hpme] and `{peacesciencer}` wants to help researchers toward that end.

>>>>

---
title: "`{peacesciencer}`: An R Package for Quantitative Peace Science Research"
thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
author:
- name: Steven V. Miller
  affiliation: Stockholm University
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
# bibliography: inst/refs.bib
abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate common types of data from scratch and to merge in ubiquitous indicators included in almost every analysis (e.g. democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at `r params$ps_website`."
keywords: "software, statistical analysis, peace science, reproducibility"
geometry: margin=1in
csl: inst/jpr.csl
mainfont: cochineal
monofont: Fira Code
#sansfont: Linux Biolinum O 
fontsize: 11pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
  ps_website: ""
  ps_manuals: ""
  ps_github: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
nocite: |
  @cartersmith2020fml, @marshalletal2017p, @coppedgeetal2020vdem, @gleditschetal2002ac, @thompsondreyer2012hir, @cowstates2016, @stinnettetal2002cow, @pevehouseetal2020tow, @schvitz2021mis, @palmeretal2021mid5, @singer1987rcwd, @nardulli2012creg, @gledtischward1999rlis, @nunnpuga2012r, @haege2011cc, @pettersson2019ov, @leedsetal2002atop, @barbierietal2009td
#always_allow_html: TRUE
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \usepackage{csquotes}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
#  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
#  - \usepackage{colortbl}
#  - \arrayrulecolor{black}
#  - \usepackage{threeparttable}
#  - \usepackage{threeparttablex}
#  - \midrule\relax
---

@@

`{peacesciencer}` addresses these problems. Built around the free and open source R programming language, `{peacesciencer}` contains a suite of data and functions for creating data of interest to researchers. Researchers can use `{peacesciencer}` to create dyad-year, leader-year, leader-dyad-year, and state-year data (among some others) from scratch. Afterward, they can add a variety of standard information (e.g. contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. `{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing conflicts. All are done with the maximum possible transparency. The project is available for public view on Github (`r params$ps_github`). The `data-raw` directory on the project's Github contains information and comments about how every data set was created. The function manuals (`r params$ps_manual`) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, `{peacesciencer}` not only assists a peace scientist with their research, but it does so in a manner that best conforms to the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

@@

`{peacesciencer}` is motivated by the following observations and ideals that led to its creation. For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's main value---if not its primary impetus---was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g. dyad-year, state-year, dispute-year) along the way that users could amend as they saw fit. `{peacesciencer}` is primarily born from this question about how to create these data from scratch. The underlying code that produces these data types is available online and `{peacesciencer}` converts these lines of code into simple functions for the ease of the researcher.

@@

tribble(~func, ~type, ~description,
        "add_archigos()", "D,S", "Add Archigos political leader information to a data frame",
        "add_atop_alliance()", "D, LD", "Add Alliance Treaty Obligations and Provisions (ATOP) alliance data to a data frame",
        "add_capital_distance()", "D, L, LD, S", "Add capital-to-capital distance to a data frame",
        "add_ccode_to_gw()", "D, L, LD, S", "Match CoW state codes to G-W state codes",
        "add_contiguity()", "D, L, LD, S", "Add CoW direct contiguity information to a data frame",
        "add_cow_alliance()", "D, LD", "Add CoW alliance data to a data frame",
        "add_cow_majors()", "D, L, LD, S", "Add CoW major power information to a data frame",
        "add_cow_mids()", "D", "Add CoW-MID data to a dyad-year data frame",
        "add_cow_trade()", "D, L, LD, S", "Add CoW trade data to a data frame",
        "add_cow_wars()", "D, S", "Add CoW war data to a dyad-year or state-year data frame",
        "add_creg_fractionalization()", "D, L, LD, S", "Add fractionalization/polarization estimates from CREG to a data frame",
        "add_democracy()", "D, L, LD, S", "Add democracy information to a data frame",
        "add_fpsim()", "D, LD", "Add dyadic foreign policy similarity measures to a data frame",
        "add_gml_mids()", "D, L, LD, S", "Add Gibler-Miller-Little (GML) MID data to a data frame",
        "add_gwcode_to_cow()", "D, L, LD, S", "Match G-W state codes to CoW state codes",
        "add_igos()", "D, S", "Add CoW intergovernmental organizations (IGO) data to a data frame",
        "add_lead()", "L, LD", "Add leader experience/attributes (LEAD) to data",
        "add_lwuf()", "L, LD", "Add leader willingness to use force estimates to data",
        "add_minimum_distance()", "D, L, LD, S", "Add minimum distance data to data frame",
        "add_nmc()", "D, L, LD, S", "Add Correlates of War National Military Capabilities data",
        "add_peace_years()", "D, S", "Add 'peace years' to dyad-year/state-year conflict data",
        "add_rugged_terrain()", "D, L, LD, S", "Add rugged terrain information to a data frame",
        "add_sdp_gdp()", "D, L, LD, S", "Add (surplus, gross) domestic product data to a data frame",
        "add_spells()", "D, L, LD, S", "Add duration 'spells' to a data frame",
        "add_strategic_rivalries()", "D, S", "Add strategic rivalry information to a data frame",
        "add_ucdp_acd()", "S", "Add Uppsala Conflict Data Program (UCDP) Armed Conflict Data to state-year data frame",
        "add_ucdp_onsets", "S", "Add UCDP onsets to state-year data",
        "filter_prd()", "D, LD", "Filter dyad-year data to just politically relevant dyads",
        "wc_duration()", "C", "Whittle Duplicate Conflict-Years by (Minimum or Maximum) Conflict Duration",
        "wc_fatality()", "C", "Whittle Duplicate Conflict-Years by Highest Fatality",
        "wc_hostility()", "C", "Whittle Duplicate Conflict-Years by Highest Hostility",
        "wc_jds()", "C", "Whittle Duplicate Conflict-Years by Just Dropping Something",
        "wc_onsets()", "C", "Whittle Uniqute Conflict Onset-Years from Conflict-Year Data",
        "wc_recip()", "C", "Whittle Duplicate Conflict-Years by Conflict Reciprocation",
        "wc_stmon()", "C", "Whittle Duplicate Conflict-Years by Lowest Start Month"
) -> add_functions

@@

`{peacesciencer}` is a package for the R programming language. This assumes at least some familiarity with the R programming language. Users should have at least version 3.5 of R, which should not be an issue since the most recent version---as of writing---is 4.1.3. `{peacesciencer}` is designed to be as user-friendly as possible. Those proficient in R, those just learning R, and those with no experience in R should be able to pick up its use fairly quickly.

@@

[^poundsign]: R users will recognize that hash signs (`#`) allow for commenting code. Users new to R should understand its use here is primarily for exposition of what the code snippets in this manuscript are doing. In subsequent code snippets, `#>` will indicate console output the user can expect from these functions.

@@

I encourage users who are using `{peacesciencer}` for the first time, especially those who are learning R for the first time because of their interest in this package, to approach `{peacesciencer}` with an idea of the kind of data they want to create for the sake of a project. The core functionality of `{peacesciencer}` begins with the creation of a data frame, which can then be populated with various indicators of interest. No matter, the suggested use of `{peacesciencer}` begins with the creation of a data frame. To start, assume a new user without much familiarity with the R programming language installed RStudio, `{tidyverse}`, and `{peacesciencer}` with the idea of using `{peacesciencer}` to help them start a new research project that seeks to explain civil conflict onset across state-years. Toward that end, they have identified their unit of analysis is state-year and can be created with the `create_stateyears()` function. To get started, I encourage entering the following command in the console in RStudio.

@@

A user who approaches `{peacesciencer}` with a project in mind will see they can better tailor this function to what they want. Their interest in a state-year analysis of civil conflict will likely gravitate them toward the Gleditsch-Ward state system data, since that is the state system that serves as the basis of the Uppsala Conflict Data Program (UCDP) armed conflict data. They will also know they have no use for pre-1946 observations since civil conflict data typically have coverage only from 1946 forward. Thus, the user can supply some additional arguments to tailor the creation of data to just what they want (here: all Gleditsch-Ward state-years from 1946 to 2019).

@@

This function has three arguments denoting the leader system to inform the creation of the leader-year data (i.e. Archigos) and two other arguments: `standardize` and `subset_years`. The `standardize` argument, which when it defaults to "none", returns all leader-years as presented in the raw Archigos data. Archigos' leader data are nominally denominated in the Gleditsch-Ward state system data, if not necessarily Gleditsch-Ward state system dates (e.g. Archigos often has leader entries prior to state system entry in a few cases). Thus, the user can standardize leader-year data to Gleditsch-Ward state system dates or CoW state system dates. Finally, the user can subset the leader-year data to a more narrow temporal domain with the `subset_years` function. If a user is interested in creating a data set of leader-years for an analysis of inter-state conflict initiation with the GML MID data, they can create the data that interests them with the following function. Notice how `create_leaderyears()` also returns information about the leader too, like the leader's approximate age that year, their gender, and information about their tenure.

@@

Researchers using `{peacesciencer}` to create data for their research project should start with one of these "create" functions, which will create the full universe of cases of interest to a researcher.

@@

Researchers should see that `{peacesciencer}`'s functionality can scale up nicely from there. For example, the following would round out the kind of information necessary to replicate Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis by adding information about national material capabilities (e.g. the composite index of national capabilities [CINC]) for both states in the dyad (`add_nmc()`), estimates of democracy for both states in the dyad (`add_democracy()`), information about alliance commitments in the dyad-year (`add_cow_alliance()`, by way of @gibler2009ima), and finishing with information about estimated population size and (surplus, gross) domestic product based on simulations reported by @andersetal2020bbgb. Whereas `add_sdp_gdp()` is the last command in the pipe-based workflow, the `{peacesciencer}` call ends by assigning to an object called `Data`. This type of assignment is done with the "right hand" assignment operator (i.e. `->`).

@@

If the user wants to move these data into Stata for analysis, they can save it to their current working directory with a command like `haven::write_dta(Data, "my-data.dta")` and import it into Stata when they are done. No matter, `{peacesciencer}` has pre-processed, cleaned, recoded, and merged the desired data that have greatly reduced the time and energy a researcher might otherwise spend doing something like hard-coding `-9s` in these data to be `NA` in the capabilities data. In this particular application, it has already created the main data required for a replication of @bremer1992ddc. There is only some slight data work to create the desired indicators for a statistical model of conflict onset, like a dummy variable for land contiguity, the presence of a major power in the dyad, and some "weak-link" indicators of militarization, relative power in the dyad, level of democracy in the dyad (using the @marquez2016qme method for extending the Unified Democracy Scores [UDS] data), and the GDP per capita in the dyad. Table \@ref(tab:tab-dd) is a formatted version of the results of a logistic regression model of conflict onset using these "dangerous dyads" indicators and temporal adjustment variables (*t*). Users typically do not end their analysis here---often looking for new predictors of conflict onset with these covariates in mind---but `{peacesciencer}` greatly reduces the time and energy researchers must invest into cleaning and processing data for analysis.

@@

`{peacesciencer}` is capable of generating data for replications of analyses at multiple levels beyond dyad-year. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with `create_stateyears(system = 'gw', subset_years = c(1946:2019))`, creating the full universe of Gleditsch-Ward state years and subsetting them to just 1946-2019 (because the UCDP data included in `{peacesciencer}` include just the observations in that time frame). Next, we can use the `add_ucdp_acd()` function to return information about ongoing UCDP conflicts and onsets for these states. `add_ucdp_acd()` takes three arguments: `type`, `issue`, and `only_wars`. `type` is an optional argument for the type of armed conflicts for which the researcher wants information. Options include "extrasystemic", "interstate", "intrastate", and "II" (short for "internationalized intrastate"). If no `type` is specified, the function returns information about ongoing disputes and onsets for all states for all types of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g. `type = c("intrastate", "II")`). `issue` is another optional argument for what issue types of conflicts the user wants. Options include "territory", "government", and "both". If no `issue` is specified, the function returns information for all conflicts regardless of the particular issue. `only_wars` is an argument that subsets the data to just those with the intensity levels of "war" when `only_wars = TRUE`. The argument defaults to `FALSE`, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, `add_ucdp_acd(type = "intrastate", only_wars = FALSE)` returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.[^adducdpacdalso]

@@

Finally, we can add some covariates of interest to these data. `add_spells()` calculates peace spells between ongoing conflicts in the data generated by `add_ucdp_acd()`. `add_democracy()` adds information about the level of democracy in the year using three prominent data sets on democracy (Polity, V-Dem, and Marquez' [-@marquez2016qme] extension of Pemstein et al's [-@pemsteinetal2010dc] UDS data). `add_creg_fractionalization()` adds information about the fractionalization and polarization of a state's ethnic and religious groups from the Composition of Religious and Ethnic Groups (CREG) Project at the University of Illinois. `add_sdp_gdp()` will add information about a state's estimated GDP, population, and GDP per capita from the @andersetal2020bbgb simulations. Finally, `add_rugged_terrain()` provide two estimates of the ruggedness of a state's terrain. The first is the terrain ruggedness index calculated by @nunnpuga2012r and the second is the @giblermiller2014etts extension of the natural logged percentage of the state that is mountainous (originally calculated by @fearonlaitin2003eicw). At the end of the pipe, the data returned by `{peacesciencer}` is assigned to an object minimally called `Data`.

`{peacesciencer}`'s tight integration with the `{tidyverse}` permits wide flexibility for the researcher. For example, assume the researcher wants to discern the estimated effect of the same set of covariates on intra-state conflicts at the threshold of war and those intra-state conflicts at or below the threshold of war. The first call included all conflicts with at least 25 deaths, per the UCDP's inclusion rules, and the peace years were calculated for those as well. If the researcher wants a new set of conflicts with a new set of peace years, it would be a simple matter of repeating the pipe-based workflow, but altering the argument in `add_ucdp_acd()` to be `only_wars = TRUE`. `{peacesciencer}` would then calculate the peace years for those (`add_spells()`). To avoid confusion with the overlapping column names, the researcher can use some `{tidyverse}` verbs to rename all those conflict variables to have a distinct prefix of `war_` (i.e. `rename_at(vars(ucdpongoing:ucdpspell), ~paste0("war_", .))`) before finally joining these data into the master data frame (i.e. `left_join(Data, .) -> Data`).  Table \@ref(tab:tab-cw) shows the fruits of the data `{peacesciencer}` generated after some post-processing and lagging important variables.

@@

`{peacesciencer}` also has support for newer levels of analysis in the peace science community, prominently leader-levels of analysis. There has been considerable emphasis in peace science research that state leaders, not "states", make foreign policy decisions that may lead to war. Understanding the attributes of leaders themselves are critical to the core research questions of the community [@ellisetal2015lead; @goemansetal2009ia; @horowitzstam2014hpme] and `{peacesciencer}` wants to help researchers toward that end.

Table \@ref(tab:tab-addfunctions) shows there are some dedicated functions for populating leader-level data with leader-specific information, in addition to adding state-year-level information (e.g. democracy, capabilities) to leader-level data.  Table \@ref(tab:tab-createfunctions) shows support for creating leader-level data that are standardized to either the CoW or G-W state system data. Suppose a researcher wanted to create leader-year data, standardized to CoW system dates, for an analysis of inter-state dispute initiation analogous to what @horowitzstam2014hpme do. The user would first start with creating the base data (`create_leaderyears(standardize = "cow", subset_years=c(1875:2010))`), which also generates some leader attributes (e.g. estimated leader age, year in office, and leader gender). From there, they would add information about leader conflict behavior in the year with `add_gml_mids()` and calculate peace-years with the `add_spells()` function.[^oninitiation] `add_lead()` will add a battery of leader experience and attributes variables from data created by @ellisetal2015lead, including whether the leader had military experience, combat experience, was a rebel fighter, and more. Finally, `add_nmc()` and `add_democracy()` will add state-year estimates of national capabilities and democracy for making state-to-state comparisons, even for leader-level analyses. Some light recoding of the data created in `{peacesciencer}` and some regression modeling reproduces the results presented in Table \@ref(tab:tab-ly), itself an approximation of the kind of leader-year analysis exemplified in @horowitzstam2014hpme.
